\documentclass[authoryear]{elsarticle}

\usepackage[T1]{fontenc}
\usepackage{ottalt}
\usepackage{xcolor}
\usepackage{mathtools}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{setspace}
\usepackage{xspace}
\usepackage{mathpartir}
\usepackage{tikz-cd}
\usepackage{comment}
\usepackage{listings}
\usepackage{MnSymbol}
\usepackage{url}
\usepackage{dashbox}
\usepackage{floatrow}
\usepackage{pifont}

\floatsetup[table]{footnoterule=none}

%\usepackage[english]{babel}

\inputott{dependent-subtyping.ott}

\newcommand{\rulehl}[2][gray!40]{
  \colorbox{#1}{$\displaystyle#2$}}
\newcommand{\redundant}[1]{\dbox{$\displaystyle#1$}}

\newcommand{\extract}[1]{\lvert #1 \rvert}
\newcommand{\lift}[1]{\lceil #1 \rceil}
\newcommand{\castup}[0]{\mathrm{cast}_\Uparrow}
\newcommand{\castdn}[0]{\mathrm{cast}_\Downarrow}
\newcommand{\dkvdash}{\vdash_\text{\tiny DK}}
\newcommand{\system}[0]{the dependent implicitly polymorphic calculus\xspace}
\newcommand{\System}[0]{The dependent implicitly polymorphic calculus\xspace}

\newcommand{\fsub}[0]{$F_{<:}$}
\newcommand{\fsubo}[0]{$???$}
\newcommand{\lami}[0]{$\lambda_{I}$\xspace}
\newcommand{\fw}[0]{$F_{\omega}$\xspace}

\newcommand{\name}[0]{$\lambda_{I}^{\forall}$\xspace}

\newcommand{\forallL}[0]{\mathbin{{\le}{\forall}} L}
\newcommand{\forallR}[0]{\mathbin{{\le}{\forall}}R}
\newcommand{\an}[3]{{\color{#2} \textsc{#1}:#3}}
\newcommand{\bruno}[1]{\an{bruno}{blue}{#1}}
\newcommand{\alvin}[1]{\an{alvin}{red}{#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\title{A Dependently Typed Calculus with Polymorphic Subtyping}

\author{Mingqi Xue}
  \ead{mqxue@cs.hku.hk}

\author{Bruno C. d. S. Oliveira}
  \ead{bruno@cs.hku.hk}

\address{The University of Hong Kong}

\begin{document}

\begin{abstract}
  A \emph{polymorphic subtyping} relation, which relates more general types
  to more specific ones, is at the core of many modern functional languages.
  As those languages start moving towards dependently typed programming a
  natural question is how can polymorphic subtyping be adapted to such settings.

  This paper presents \system (\name): a simple dependently typed calculus
  with polymorphic subtyping. The subtyping relation in \name
  generalizes the well-known polymorphic subtyping relation by
  Odersky and L\"aufer (1996). Because \name is dependently typed,
  integrating subtyping in the calculus is non-trivial. To overcome
  many of the issues arising from integrating subtyping with dependent
  types, the calculus employs \emph{unified subtyping}, which is a technique
  that unifies typing and subtyping into a single relation. Moreover, \name
  employs explicit casts instead of a conversion rule, allowing
  unrestricted recursion to be naturally supported.
  We prove various non-trivial results, including \emph{type soundness} and \emph{transitivity}
  of unified subtyping. \name and all corresponding proofs
  are mechanized in the Coq theorem prover.
  %and we also have a simple prototype
  %implementation of \name.
%A complication that arizes in the type-soundness proof
%is that a direct semantics is non-deterministic due
\end{abstract}


\begin{keyword}
  Type Systems \sep Dependent Types \sep Subtyping \sep Polymorphism
\end{keyword}
% \keywords{Type System, Dependent Type, Subtyping, Polymorphism}

\maketitle

\input{introduction.filt}
\input{overview.filt}
\input{type-system.filt}
\input{metatheory.filt}
\input{discussion.filt}
\input{related-work.filt}
\input{conclusion.filt}

% \bibliographystyle{plainurl}
\bibliographystyle{elsarticle-num-names}
\bibliography{reference,RW,RW2}

\end{document}
